Analysis of observational data, to calculate the Delayed-Tau model parameters

### 2025-03-16 thesis.qmd
cmd-ms comparison

-   $\tau \sim \text{uniform(lower bound = 1 Gyr, upper bound = 13.8 Gyr)}$,
-   $t_{sf}\sim \text{uniform(lower bound = 1 Gyr, upper bound = 13.8 Gyr)}$,
-   $\zeta \sim \text{normal(mean=1.3, }\sigma = 0.01)$, within the limits $[1,2]$,

and the likelihood $\log(\text{SFR}_{obs}) \sim \text{normal}\left(\log(\text{SFR}_\text{MCMC}), \sigma = 0.1\right)$, while taking the Stellar Masses of the galaxies as they are, without any variance, and $A_{del}$ is calculated by @eq-A. The MCMC run consists of 6 chains with 8000 iterations each (2500 iterations for the burn-in phase). Using this method, the galaxies can have different $t_{sf}$, while only taking positive values for their timescales.

The prior of the $\tau$ and $t_{sf}$ follow a uniform distribution, since we do not want to introduce any unwanted bias in the value selection of the MCMC sampler. While there is bibliography that could justify using a normal distribution for $\tau$, for example @speagleHighlyConsistentFramework2014 and @kroupaConstraintsStarFormation2020 explain that the galaxies of the Main Sequence are expected to have $3.5\lesssim\tau/\text{Gyr}\lesssim4.5$, this could significantly influence the results of galaxies with slow rising SFR's. In the same way a wrong prior for $t_{sf}$ could effect the results, especially of younger galaxies. While the choice of different priors is possible, for the aim of this thesis, the uniform distribution is the best choice.

@@

![Comparison of the observed SFR with the predicted by the parameters given by the MCMC run.](r_mcmc/plots/sfr_diff_plot.png){#fig-comp-sfr-obs-pred}

![Linear fit for $A_{del}\sim M_*$, with a correlation of 94%. The black dashed line is the linear regression weighted by $x$, with a fit $\log(A_{del}) = 0.95\cdot\log(M_*)+0.60, with a corralation of 98\%$](method_comparison/A_M_MCMC.png){#fig-A-M-mcmc}

@@

![$A_{del}-\text{sSFR}$ plot, where $A_{del}$ is calculated with different methods. In blue is the normalization constant as it is calculated by the MCMC method, while in red are results from Newton-Raphson.](method_comparison/sSFR_A.png){#fig-sSFR-A-both-methods}

>>>>

-   $t_{sf}\sim \text{uniform(lower bound = 1 Gyr, upper bound = 13.8 Gyr)}$,
-   $\zeta \sim \text{normal(mean=1.3, }\sigma = 0.01)$, within the limits $[1,2]$,

and the likelihood $\log(\text{SFR}_{obs}) \sim \text{normal}\left(\log(\text{SFR}_\text{MCMC}), \sigma = 0.1\right)$, while taking the Stellar Masses of the galaxies as they are, without any variance, and $A_{del}$ is calculated by @eq-A. For the $\tau$ we will consider two different priors, $\tau \sim \text{uniform(lower bound = 1 Gyr, upper bound = 13.8 Gyr)}$ and $\tau \sim \text{normal(mean = 4 Gyr, }\sigma  = 13.8 \text{Gyr)}$. The MCMC run consists of 6 chains with 8000 iterations each (2500 iterations for the burn-in phase). Using this method, the galaxies can have different $t_{sf}$, while only taking positive values for their timescales.

The prior of the $\tau$ and $t_{sf}$ follow a uniform distribution, since we do not want to introduce any unwanted bias in the value selection of the MCMC sampler. However, there is bibliography that could justify using a normal distribution for $\tau$, for example @speagleHighlyConsistentFramework2014 and @kroupaConstraintsStarFormation2020 explain that the galaxies of the Main Sequence are expected to have $3.5\lesssim\tau/\text{Gyr}\lesssim4.5$. The same prior will not apply for $t_{sf}$ since it could significatly effect the results, especially of younger galaxies.

@@

::: {#fig-comp-sfr-obs-pred layout-ncol="2"}

![MCMC run for $\tau \sim \text{uniform}$ prior.](r_mcmc_uni/plots/sfr_diff_plot.png){#fig-comp-sfr-obs-pred-uni}

![MCMC run $\tau \sim \text{normal}$ prior.](r_mcmc_normal/plots/sfr_diff_plot.png){#fig-comp-sfr-obs-pred-norm}

Comparison of the observed SFR with the predicted by the parameters given by the MCMC run.
:::

::: {#fig-A-M-mcmc layout-nrow="2"}

![Linear fit for $A_{del}\sim M_*$, with a correlation of 94%. The black dashed line is the linear regression weighted by $x$, with a fit $\log(A_{del}) = 0.95\cdot\log(M_*)+0.60, with a corralation of 98\%$](method_comparison/A_M_MCMC_uni.png){#fig-A-M-mcmc-uni}

![Linear fit for $A_{del}\sim M_*$, with a correlation of 99%. The black dashed line is the linear regression weighted by $x$, with a fit $\log(A_{del}) = 0.97\cdot\log(M_*)+0.44, with a corralation of 99.5\%$](method_comparison/A_M_MCMC_norm.png){#fig-A-M-mcmc-norm}
:::

@@

![$A_{del}-\text{sSFR}$ plot, where $A_{del}$ is calculated with different methods. In blue and green are the normalization constant values as they calculated by the MCMC method with different priors, while in red are results from Newton-Raphson.](method_comparison/sSFR_A.png){#fig-sSFR-A-both-methods}
